---
title: "visualization-analysis"
output: html_document
date: "2025-01-05"
data-set-source: https://www.kaggle.com/datasets/crawford/80-cereals/data
---

*this document is for practicing visualizations*

```{r message=FALSE, warning=FALSE}
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  corrr, corrplot,# correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  ggplot2,hrbrthemes       # plots
  
)
```

importing table and initial understanding of the table

```{r}
cereal <- import("cereal.csv") 
```

mfr: Manufacturer of cereal
    A = American Home Food Products;
    G = General Mills
    K = Kelloggs
    N = Nabisco
    P = Post
    Q = Quaker Oats
    R = Ralston Purina 

check out document for more break downs


```{r}
skim(cereal)
```
```{r}
head(cereal)
```
- no missing values
- 3 categorical, 13 numeric (nutritional values)

### pie charts

```{r simple pie chart of manufacturer}
pie(table(cereal$mfr), main="Pie Chart of Manufacturer")
```

pie charts are hard to read so let's move to bar plots

guide used: https://r-graph-gallery.com/218-basic-barplots-with-ggplot2.html

### bar plots

```{r simple bar plot of frequency}
ggplot(cereal) +
  geom_bar(aes(x=mfr)) +
  labs(y = "Count", x = "Manufacturer")
```

### controlling the color 

ggplot(cereal, aes(x=as.factor(mfr))) +
geom_bar(color="blue)

```{r}
ggplot(cereal, aes(x=as.factor(mfr))) + 
  geom_bar(color="blue", fill="dodgerblue") + 
  labs(y = "Count", x = "Manufacturer")
```

can flip with coord_flip

```{r}
ggplot(cereal, aes(x=as.factor(mfr))) + 
  geom_bar(color="blue", fill="dodgerblue") + 
  labs(y = "Count", x = "Manufacturer") +
  coord_flip()
```
changing the width of the bars

```{r}
ggplot(cereal, aes(x=as.factor(mfr))) + 
  geom_bar(color="blue", fill="dodgerblue", width = 0.2) + 
  labs(y = "Count", x = "Manufacturer") 
```

### reorder a bariable with ggplot2
https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html

using forcats library

```{r reordering barplot}
ggplot(cereal) +
    geom_bar(aes(x = forcats::fct_infreq(mfr),
                 fill=type)) + 
  labs(y = "Count", x = "Manufacturer") 
```

Boxplots of Nutritional Values

```{r boxplot with base r}
boxplot(rating~mfr, data=cereal)
```

boxplot (can use violinplot because there are only 1-2 points for mfr:A)

```{r}
cereal %>% 
  ggplot(aes(x=mfr, y=rating, fill=mfr)) +
  geom_boxplot() +
  guides(fill="none")
```

```{r highlight boxplot}
cereal %>% 
  mutate( type=ifelse(mfr=="Q","Highlighted","Normal")) %>% 
  ggplot( aes(x=mfr, y=rating, fill=type, alpha=type)) + 
  geom_boxplot() +
    scale_fill_manual(values=c("#69b3a2", "grey")) +
    scale_alpha_manual(values=c(1,0.1)) +
    theme_ipsum() +
    theme(legend.position = "none") +
    xlab("")
```

```{r correlation}
corr_columns <- cereal %>% select(c(calories:vitamins))

corrr::correlate(corr_columns)
```


```{r correlation table with color}
M<-cor(corr_columns)
corrplot(M)
```
Nothing really stands out like potassium and fiber, and calories and sugar

### testing different corrplots
```{r}
corrplot(M, method="pie")
```


```{r color correlation table}
corrplot(M, method="color")
```

```{r color number}
corrplot(M, method="number")
```






Questions

By Manufactorer what is the average protein, calories, etc?

```{r}
cereal %>% 
  group_by(mfr) %>% 
  dplyr::summarise(mean = mean(calories), n=n())
```






What cereal has the best rating?







